---
output: github_document
---

# US building permit

```{r setup}
library(tidyverse)
```

## Data import

Data downloaded from <https://www.recenter.tamu.edu/data/building-permits/>

```{r}
permits_raw <- read_csv("dataPermit_full.csv", na = "null")
permits_raw
```

* area = metropolitan standard area
* date = month / year (character vector, or strings)
* f1 = 1 family house, f24 = 2-4 familes, f5 = 5+ family house
* units = number of buildings, change in units; value = average value of building; valchange = change in value

```{r}
permits <- permits_raw %>%
  separate(date, c("month", "year"), "/", convert = TRUE) %>%
  filter(year > 2007) %>%
  select(-ends_with("change"))
permits
```

## Basic eda

```{r}
permits %>% count(year)
permits %>% count(area)
permits %>% count(area) %>% count(n)
```

I'm focus on single family homes

```{r}
permits <- permits %>% 
  mutate(date = year + (month - 1) / 12)

ggplot(permits, aes(date, f1units)) + 
  geom_line(aes(group = area))
```

## Focus

Focus on big cities. Possible bias our results, because big cities are small cities? 
```{r}
f1units <- permits %>% 
  group_by(area) %>%
  summarise(mean = mean(f1units)) %>% 
  arrange(desc(mean))
f1units

f1units %>% filter(mean > 100)

permits_big <- permits %>%
  semi_join(f1units %>% filter(mean > 100))
permits_big
```


```{r}
ggplot(permits_big, aes(date, f1units)) + 
  geom_line(aes(group = area))

ggplot(permits_big, aes(date, f1units)) + 
  geom_line(aes(group = area), alpha = 1/10)

ggplot(permits_big, aes(date, f1units)) + 
  geom_line(aes(group = area), alpha = 1/10) + 
  scale_y_log10()

ggplot(permits_big, aes(date, f1units)) + 
  geom_line(aes(group = area), alpha = 1/10) + 
  scale_y_log10() + 
  geom_smooth(se = FALSE)

```

## Model

```{r}
houston <- permits %>% filter(str_detect(area, "Houston"))
houston %>% count(area)

ggplot(houston, aes(date, f1units)) + 
  geom_line(aes(group = area))

ggplot(houston, aes(month, f1units)) + 
  geom_line(aes(group = year)) +
  scale_y_log10()

```

Question:

* Is this pattern the same everywhere?
* What drives it? Is it the weather?
* Houston in July is less pleasant than Houston Dec

```{r}
library(modelr)


houston_mod <- lm(log(f1units) ~ factor(month), data = houston)

houston %>% 
  add_predictions(houston_mod) %>%
  ggplot(aes(date, pred)) + 
  geom_line()

houston %>% 
  add_residuals(houston_mod) %>%
  ggplot(aes(date, resid)) + 
  geom_line()
  
```

* What's driving this trend?
* What hapened around 2010?

## Extend the model to every city

```{r}
by_area <- permits_big %>%
  group_by(area) %>%
  nest()

area_model <- function(df) {
  lm(log10(f1units + 1) ~ factor(month), data = df)
}

detrended <- by_area %>% mutate(
  model = map(data, area_model),
  resids = map2(data, model, add_residuals)
) %>% unnest(resids)

ggplot(detrended, aes(date, resid)) + 
  geom_line(aes(group = area), alpha = 1/10) + 
  geom_smooth(se = FALSE)

```

