# US building permits

```{r setup}
library(tidyverse)
```

## Data import

Data downloaded from <https://www.recenter.tamu.edu/data/building-permits/>

```{r}
permits_raw <- read_csv("dataPermit_full.csv", na = "null")
permits_raw
```

* f1 = 1 family home; f24 = 2-4 family home; f5 = 5+ unit home
* units = number built; change = percent change from previous month
* value = average value per unit; valchange = percent change from previous month
Hire Director of Education

Eliminate the change and valchange variables; can easily recompute if needed (and good practice to be sceptical about someone elses computations)

```{r}
permits <- permits_raw %>%
  select(-ends_with("change")) %>%
  separate(date, c("month", "year"), "/", convert = TRUE) %>%
  filter(year > 2007)
permits
```


## Very basic EDA

```{r}
permits %>% count(year)
permits %>% count(area)
permits %>% count(area) %>% count(n)
```

I'm going to arbitrarily focus on single family houses. 

```{r}
ggplot(permits, aes(year + month / 12, f1units)) + 
  geom_line(aes(group = area))
```

## Focus on big cities

TOO MUCH DATA.

Let's start by focusing on the bigger cities. This might give a misleading overview (maybe big cities are different fundamentally) but it's a good place to start because they'll have more permits so should hopefully be less variation and the long-term patterns clearler.

```{r}
f1units <- permits %>% 
  group_by(area) %>%
  summarise(mean = mean(f1units)) %>% 
  arrange(desc(mean))
f1units

permits_big <- permits %>% 
  semi_join(f1units %>% filter(mean > 120)) %>%
  mutate(date = year + (month - 1) / 12)
```

Now replot:

```{r}
ggplot(permits_big, aes(date, f1units)) + 
  geom_line(aes(group = area))
```

```{r}
ggplot(permits_big, aes(date, f1units)) + 
  geom_line(aes(group = area)) + 
  scale_y_log10()

ggplot(permits_big, aes(date, f1units)) + 
  geom_line(aes(group = area), alpha = 1/10) + 
  scale_y_log10()

ggplot(permits_big, aes(date, f1units)) + 
  geom_line(aes(group = area), alpha = 1/5) + 
  scale_y_log10() + 
  geom_smooth(se = FALSE)
```

## What's that regular pattern?

```{r}
houston <- permits_big %>% filter(str_detect(area, "Houston"))
ggplot(houston, aes(date, f1units + 1)) +
  geom_line()

ggplot(houston, aes(month, f1units + 1)) +
  geom_line(aes(group = year)) +
  geom_smooth(se = FALSE)

ggplot(houston, aes(month, f1units + 1)) +
  geom_line(aes(group = year)) +
  geom_smooth(se = FALSE) + 
  scale_y_log10()

```

Wonderings:

* Is the seaonal pattern more pronounced in colder climates?
* What's the lag between permitting and beginning construction?
* What drives seasonal pattern? July usually less pleasant than Jan in Houston.

## Can we remove it?

Let's use a model to partition the signal into monthly pattern + other

```{r}
library(modelr)

houston_mod <- lm(log(f1units) ~ factor(month), data = houston)
houston %>%
  add_predictions(houston_mod) %>%
  ggplot(aes(date, pred)) +
  geom_line()

houston %>%
  add_residuals(houston_mod) %>%
  ggplot(aes(date, resid)) +
  geom_hline(yintercept = 0, colour = "white", size = 3) +
  geom_line()
```

## Now for all cities

```{r}
by_area <- permits_big %>%
  group_by(area) %>%
  nest()

area_model <- function(df) {
  lm(log10(f1units + 1) ~ factor(month), data = df)
}

detrended <- by_area %>% mutate(
  model = map(data, area_model),
  resids = map2(data, model, add_residuals)
) %>% unnest(resids)


ggplot(detrended, aes(date, resid)) +
  geom_line(aes(group = area), alpha = 1/10) +
  geom_smooth(se = FALSE)

ggplot(detrended, aes(date, 10 ^ resid)) +
  geom_line(aes(group = area), alpha = 1/10) +
  geom_smooth(se = FALSE)

ggplot(detrended, aes(date, 10 ^ resid)) +
  geom_line(aes(group = area), alpha = 1/10) +
  stat_summary(geom = "line", fun.y = function(x) quantile(x, 0.25), colour = "blue") +
  stat_summary(geom = "line", fun.y = function(x) quantile(x, 0.75), colour = "blue")
```

Wonderings:

* Is the gap spreading?
* Can explain why some cities are growing more?
* Is it population? Is it geography?  (Probably more complex than either!)
* Does the pattern follow for 2-4 and 5+ family units?
